---
title: "Who are the top 5 most valuable soccer players, what is their nationality, position, team (squad), and how many goals did they score in the 2018/2019 season?"
output: html_document
---

```{r setup, include = FALSE, echo = FALSE}
library(tidyverse)
library(lubridate)
library(gt)
library(paletteer)
library(plotly)
library(flexdashboard)
library(janitor)
library(forcats)
library(purrr)
library(webshot)
library(graphics)
library(RColorBrewer)
library(DataExplorer)
library(magrittr)
library(sf)
library(leaflet)
library(ggridges)
library(tigris)
library(ggthemes)
library(htmltools)
library(htmlwidgets)
library(readr)

#Working directory for .RMD
knitr::opts_knit$set(echo = TRUE,
                     root.dir = rprojroot::find_rstudio_root_file())

#Controlling figure output in markdown
knitr::opts_chunk$set(
  fig.height = 12,  
  fig.width = 8,
#  fig.asp = .5,
  out.width = "90%",
#  out.height = 
 fig.align  = "center",
  cache = FALSE,
  eval  = TRUE,
  echo  = TRUE,
  warning = FALSE
)

#My Colours (from viridis)
my_purple <- "#440154FF"
my_yellow <- "#FDE725FF"

#Set Theme for ggplot2
theme_set(theme_bw() + theme(plot.title = element_text(hjust = 0.5,face = "bold"), legend.position = "bottom"))

#Set Scientific notation output and decimal places for knitr
options(scipen = 999)
options(digits = 4)
options(dplyr.summarise.inform = FALSE)
```


```{r Read, wrap and clean data, include = FALSE}

# Load data

# Load soccer stats data
soccer.df <- read_csv(file = './data/soccer.csv') %>%
  filter(!is.na(player))

# Load country names data set (with abbreviations)
ioc_country.df <- read_csv(url("https://raw.githubusercontent.com/datasets/country-codes/master/data/country-codes.csv")) %>%
  select(FIFA,official_name_en) %>%
  rename(Country=official_name_en,NOC=FIFA)

# Load country coordinates (geometry) data set
world_boundaries3.df <- read_csv("./data/world-coordinates-data.csv") %>%
janitor::clean_names() 


```



```{r Data wrangling, echo = FALSE, include = FALSE}
# Do data wrangling and create needed variables
soccer.df <- soccer.df %>%
  mutate(value_millions = (value/1000000)) %>%
  separate(nationality, into = c("nationality", "country_origin"), sep = " ")

# Create variable ranking of players by value and goals
soccer.df <- soccer.df %>%
  arrange(desc(value)) %>%
  mutate(ranking_player_value = 1:length(player)) %>%
  arrange(desc(goals)) %>%
  mutate(ranking_player_goals = 1:length(player))

# Separating geometry into latitude and longitude
world_boundaries3.df <- world_boundaries3.df %>%
  separate(geo_point_coord, into = c("lat", "long"), sep = ",") %>%
  mutate(lat = as.numeric(lat),long = as.numeric(long))


# Join country's names
soccer_world.df <- left_join(
  soccer.df,
  ioc_country.df,
  by = c("country_origin" = "NOC") #zipcode in first, zip_code second
)


# Edit names for countries from United Kingdom, and other countries with special characters

soccer_world.df <- soccer_world.df %>%
  mutate(Country = ifelse(country_origin %in% "ENG"|country_origin %in% "WAL"|country_origin %in% "SCO"|country_origin %in% "NIR","United Kingdom",Country)) %>%
  mutate(Country = ifelse(country_origin %in% "CUW","Curaçao",Country)) %>%
  mutate(Country = ifelse(country_origin %in% "KVX","Kosovo",Country)) %>%
  mutate(Country = ifelse(country_origin %in% "CIV","Ivory Coast",Country)) %>%
  mutate(Country = ifelse(country_origin %in% "CUW","Curacao",Country)) %>%
  mutate(Country = ifelse(nationality %in% "GPE","Curacao",Country)) %>%
  mutate(Country = ifelse(country_origin %in% "REU","Reunion",Country))


# Joining soccer_world.df with world boundaries data

soccer_world3.df <- inner_join(
  soccer_world.df, 
  world_boundaries3.df, 
  by = c("Country" = "english_name") 
)

```

 
</font>

-**Table 1** shows the country of origin, goals, assists of the 5 most valuable players by their position in the 2018/2019 season. The value of forwards is numerically higher than the value of midfielders, goakeepers or defenders. As a matter of fact, our visualization suggests the presence of an association between the player's value and the number of goals.

-**Figure 1** shows the 5 players with the highest value in the top 5 European leagues in an interactive visualization. A large proportion of these players were strikers and showed a high number of goals and assists during the analyzed season.



```{r, echo=FALSE}
# Plot most valuable players by team
soccer_world.df %>%
  mutate(Position = ifelse(grepl("Forward", position2),"Forward",position2)) %>%
  mutate(Position = ifelse(grepl("Midfielder", Position),"Midfielder",Position)) %>%
  mutate(Position = ifelse(grepl("Defender", Position),"Defender",Position)) %>%
  group_by(Position) %>%
  arrange(ranking_player_value) %>%
  slice(1:5) %>%
  mutate(average_value_top_5 = round((mean(value,na.rm = TRUE)),2)) %>%
  arrange(desc(average_value_top_5)) %>%
  select(c(player,squad,Country,Position,goals,assists,value)) %>%
  rename("Player" = player,"Team" = squad,"Goals" = goals,"Assists" = assists,"Value (USD)" = value) %>%
  gt() %>%
  fmt_number(
    columns = where(is.numeric),
    decimals = 0 ,
    suffixing = TRUE) %>%
  cols_align(
    align = "center",
    columns = everything()) %>%
  data_color(
    columns =Goals,
    colors = scales::col_numeric(
      palette = c("PuBu"),
      domain  = c(min(Goals), max(Goals)))) %>%
  data_color(
    columns =Assists,
    colors = scales::col_numeric(
      palette = c("PuBu"),
      domain  = c(min(Assists), max(Assists)))) %>%
  data_color(
    columns =`Value (USD)`,
    colors = scales::col_numeric(
      palette = c("Purples"),
      domain  = c(min(`Value (USD)`), max(`Value (USD)`)))) %>%
  summary_rows(
    groups  = TRUE,
    columns = where(is.numeric),
    fns = list("# Average top 5" = ~mean(., na.rm = TRUE)),
    formatter = fmt_number,
    decimals  = 0,
    suffixing = TRUE) %>%
  tab_header(title = md("**Table 1. Top 5 most valuable players by position in 2018/2019 season**")) %>%
  tab_source_note("*Data merged from transfermarkt.de and fbref.com by Rafael Stepien and available on Kaggle*")

```

***

```{r, echo=FALSE}
# Plot most valuable players by league

Plot_valuable_players <- soccer_world.df %>%
  mutate(Position = ifelse(grepl("Forward", position2),"Forward",position2)) %>%
  mutate(Position = ifelse(grepl("Midfielder", Position),"Midfielder",Position)) %>%
  mutate(Position = ifelse(grepl("Defender", Position),"Defender",Position)) %>%
  group_by(league) %>%
  arrange(ranking_player_value) %>%
  slice(1:5) %>%
  mutate(average_value_top_5 = round((mean(value,na.rm = TRUE)),2)) %>%
  arrange(desc(average_value_top_5)) %>%
  mutate(
    text_label = str_c("Value, Nationality, and Squad and More",
                       "\nTeam: ", squad,
                       "\nValue: ", value_millions,
                       "\nGoals: ", goals,
                       "\nAssists: ", assists,
                       "\nPosition: ", position2)) %>%
  rename("Player" = player,"Team" = squad,"Value" = value_millions,"Goals" = goals) %>%
  ggplot() +
  geom_col(aes(x = Value, y = fct_reorder(Player,value), fill = fct_reorder(Position,value,.desc=TRUE), text = text_label)) +
  labs(
    x     = "Value (millions USD)",
    y     = "Player",
    title = "Figure 1. Most valuable players in top 5 European Leagues in 2018/2019 season" ) +
  scale_fill_viridis_d("Position") +
  theme(axis.text.x = element_text(angle = 0,
                                   vjust = 0.5,
                                   hjust = 1,
                                   size  = 8)) +
  facet_wrap(~league , scales = "free", ncol = 2) + theme(panel.spacing = unit(7, "lines"))

# Transform into a ggplotly object
ggplotly(Plot_valuable_players, tooltip = "text") %>% layout(width = 900,heights = 1000,margin = list(b = 90))
```

